### 2018-05-16 Project/projthought.tex
First Commit

\documentclass{article}
\usepackage{amsmath}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{listings}
\usepackage{float}
\geometry{left = 1in, right =1in,top=1in, bottom = 1in}
\doublespacing
\begin{document}
\newcommand*{\be}{\mathbb{E}}
\newcommand*{\bv}{\mathbb{V}}
\lstset{showspaces = false, showstringspaces = false}
\linespread{2}
\title{Project Thoughts}
\author{Donald DiJacklin}
\maketitle
Profit from breeding snakes depends on the revenue from either selling the snake this period or breeding it this period and selling its babies next period and the costs of feeding it

\end{document}